\documentclass[a4paper,11pt]{article}
\usepackage{subfiles}
\usepackage{modules/f420}

\title{Geodesics, Shortest Paths, and Non-Isomorphic Nets}
\author{{\scshape Potvin} Nicolas \and {\scshape Zolotov} Boris}
\date{ULB,\quad November 2020}

\begin{document} \maketitle

\section{Introduction}

Given a collection of 2D polygons, a \emph{gluing} or a \emph{net} describes a closed surface by specifying how to glue (a part of) each edge of these polygons onto (a part of) another edge. Alexandrov's uniqueness theorem~\cite{alex} states that any valid gluing that is homeomorphic to a sphere and that does not yield a total facial angle greater than $2\pi$ at any point, corresponds to the surface of a unique convex 3D polyhedron (doubly covered convex polygons are also regarded as polyhedra). Note that the original polygonal pieces might need to be folded to obtain this 3D surface.

Unfortunately, the proof of Alexandrov's theorem is highly non-constructive. The only known approximation algorithm to find the vertices of this polyhedron~\cite{kpd09-approx} has a (pseudopolynomial) running time really large in $n$, where $n$ is the total complexity of the gluing.

There is no known exact algorithm for reconstructing the 3D polyhedron, and in fact the coordinates of the vertices of the polyhedron might not even be expressible as a closed formula~\cite{bannister2014galois}.

Enumerating all possible valid gluings is also not an easy task, as the number of gluings can be exponential even for a single polygon~\cite{DDLO02}. However one valid gluing can be found in polynomial time using dynamic programming~\cite{DO07,lo96-dynprog}. Complete enumerations of gluings and the resulting polyhedra are only known for very specific cases such as the Latin cross~\cite{ddlop99} and a single regular convex polygon~\cite{DO07}.

The special case when the polygons to be glued together are all identical regular $k$-gons, and the gluing is \emph{edge-to-edge} was studied recently for $k \ge 6$~\cite{kl17-hex} and $k=5$~\cite{alz-penta}. The aim of this project is to study the case of $k=4$.

\input{NicolasPart}

\section{Boris' part: enumeration of gluings, isomorphism check}

Enumeration of edge-to-edge gluings was already done for hexagons and pentagons
in~\cite{kl17-hex,alz-penta}. This time we also want the running time of the enumeration
procedure to be polynomial in the number of polygons, not exponential. This seems
possible to do since the number of faces of the polyhedron we can get is bounded,
and there is polynomial number of options to draw each face on the grid.

Due to Gauss—Bonnet formula we know that the number of vertices of a polyhedron
glued edge-to-edge from squares is at most 8. We then want to clusterize isomorphic
polyhedra. One way is to see if they are homothetic, for this we can apply Chen—Han's
algorithm: take the matrix of distances it calculates and see if we can transform
one matrix into another by permutations of rows and columns and multiplying
the entire matrix by a constant. Since all the matrices are at most $8 \times 8$,
this is a fast check.

In the end we will be able to see if all possible graph structures of polyhedra
with at most 8 vertices are realizable by gluing from squares.

\input{APISpecs}

\bibliography{boris-bac}{}
\bibliographystyle{plain}

\end{document}
